---
title: "An Exporatory Vignette for sigex and sigexUI"
author: Andrew, Jim, ...
output: 
  pdf_document:
    highlight: default
    number_sections: true
    toc: true
    extra_dependencies:
      common: null
date: "08/26/2021"
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{An Exporatory Vignette for sigex and sigexUI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  prompt = TRUE,
  comment = ""
)
```

```{r setup, include = FALSE}
library(sigex)
library(sigexUI)
set.seed(1235)
```

\newpage

# Introduction
\label{sec:intro}
For now, let's just make some notes about the important data structures in `sigex`. These are initially just unconfirmed notes from Andrew which may not be correct.

Recall that $N$ is the number of variables in a multivariate time series, and $T$ is the number of time points observed in the series. 
A time series is decomposed into $K$ latent components
<!-- -->
\begin{align*}
\vec{Y}_t = \sum_{k=1}^K \vec{S}_t^{(k)},
\end{align*}
<!-- -->
and a filtered series is written as
<!-- -->
\begin{align*}
\delta(B) \vec{Y}_t = \sum_{k=1}^K \delta(B)^{(-k)} \underline{\vec{S}}_t^{(k)}.
\end{align*}

# Data Structures
\label{sec:data-structures}

## Constraint
A matrix `cbind(b, C)` representing the constraint $\vec{C} \vec{\psi} = \vec{b}$. Can also be `NULL` to represent no constraints.

## Generalized Cholesky Decomposition (gcd)
An unlabeled list of two elements:

1. A matrix $\vec{L}$,
2. A vector representing the elements of a diagonal matrix $\vec{D}$,

such that the given matrix $\vec{A}$ is decomposed into $\vec{L} \vec{D} \vec{L}^\top$.

## Model (mdl)
A named list of elements:

1. \textbf{ranks}: a numbered list of length $K$. The $k$ element of the list is "a vector of integers between $1$ and $N$, corresponding to indices of non-zero Schur complements in the GCD of the innovations' covariance matrix for the new latent component".

2. \textbf{type}: a numbered list of length $K$. The $k$th element is another numbered list with four entries:

	1. A string which represents the model type of the component.
	2. A vector which represents the order of the component. For example,
	   an $\text{ARMA}(p,q)$ component will have the vector $c(p,q)$ here.
	3. A vector of four numbers which represent "bounds for `rho` and
	   `omega`, the cycle parameters of the new latent component: `rho` lies
	   in `(bounds[1], bounds[2])` and `omega` lies in `(bounds[3], bounds[4])`".
	4. A string which represents the epithet of the component.

3. \textbf{diffop}: a numbered list of length $K$. The $k$th element is a vector of coefficients representing $\delta^{(k)}(B)$.

4. \textbf{regress}: a numbered list of length $N$ where the $i$th element is a $T$ dimensional vector $x_{i1}, \ldots, x_{iT}$ representing a regressor series.

## Parameter (par)
A unlabeled list with four elements:

1. A list of $K$ elements, where the $k$th element is an $N \times N$ lower-triangular matrix $\vec{L}_k$ for a covariance matrix.
2. A list of $K$ elements, where the $k$th element is an $N$ dimension vector which represents the logarithm of diagonal elements of $\vec{D}_k$  for a covariance matrix.
3. A list of $K$ elements, where the $k$th element contains time series parameters for the $k$th latent process. For example, a component which is an $\text{ARMA}(p,q)$ process will have an $N \times N \times (p+q)$ array here.
4. An $N$ dimensional vector corresponding to regression parameters.

The matrix $\vec{L}_k \vec{D}_k \vec{L}_k^\top$ represents the covariance $\vec{\Sigma}_k$ for the $k$th latent component of the model.

## MLE Output (mlefit)
An unlabeled list with two elements:

1. The output of the optimizer: `optim` or `nlminb`
2. The estimates in "par" parameterization.

## MOM Output (momfit)
The estimates in "par" parameterization.

## Psi
TBD

## Pre-parameter (pre)
TBD

# References

