---
title: "sigexUI-overview"
author: "AR & JL"
date: "4/19/2022"
output: 
  ioslides_presentation:
    smaller: true
    hitheme: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

################################################################################
## Disclaimer

This article is released to inform interested parties of ongoing research and to
encourage discussion of work in progress.
Any views expressed are those of the authors and not
those of the U.S. Census Bureau.


################################################################################
## Why?

Following feedback from JSS and other users we aim to make `sigex` fit more
seamlessly into the R ecosystem.

- User friendly defaults
- Error/wanrining checks
- Esstablished R function support
	* `residuals()`
	* `summary()`
	* `print()`

Want to "future proof" development of future add-ons.

- Don't make classes or helper functions too rigid
- Think about the high level structures of multivariate unobserved component models


################################################################################
## Class system in R


################################################################################
## Example: fitting ARIMA model

We have all fit an ARIMA model to the AirPassengers Series. 

```{r, echo = TRUE}
fit <- arima(AirPassengers, order = c(1, 0, 0))
class(fit)
```

################################################################################
## Example: fitting ARIMA model

The R ecosystem and its users is now familiar with the concepts of `fit` being
the output of a model and with that comes ceritain expectations about how it should 
behave. For example,

```{r, collapse=TRUE}
coef(fit)
```
```{r, collapse=TRUE}
AIC(fit)
```
```{r, collapse=TRUE}
logLik(fit)
```
```{r, collapse=TRUE}
predict(fit, n.ahead = 2)
```

################################################################################
## Example: fitting ARIMA model

How does this happen?

```{r}
str(fit)
```


################################################################################
## Example: fitting ARIMA model

Let's look at the source code for *arima*

<https://github.com/SurajGupta/r-source/blob/master/src/library/stats/R/arima.R>



################################################################################
## Some Definitions

- **Classes** user-defined data types, blueprint for individual objects, attributes and methods.
- **Object** instances of a class created with specifically defined data. Objects can correspond to real-world objects or an abstract entity.
- **Methods**
- **Generic Function**
- **Method Dispatch**



